Spatiotemporal Multi-Resolution Approximation of the Amari Type Neural Field Model

### 2012-01-19 Multiresolution_framework/ltx/manuscript.tex
added some paragraphs to estimation section

\singlespacing
\begin{table*}[!t]
\begin{tabular}{|l|l|l|}
	\hline
	\textbf{Symbol} & \textbf{Quantity} & \textbf{Units}\\
	\hline
	\multicolumn{3}{|c|}{\emph{Domain and indices}}\\
	\hline
	$\Omega$ & spatial domain & n.a.\\
	$\mathbf{r}$ & spatial location & [mm, mm]\\
	$t$ & time & s\\
	\hline
	\multicolumn{3}{|c|}{\emph{Model}}\\
	\hline
    $y_t$ & observation & mV\\
    $v(\mathbf{r},t)$ & mean membrane potential & mV \\
	$f(v\left(\mathbf{r},t\right))$ & firing rate function & spikes.s$^{-1}$\\
	$e(\mathbf{r},t)$ & field disturbance, with covariance function $\gamma(\mathbf r-\mathbf r')$ & mV\\
	$\boldsymbol\varepsilon_t$ & observation noise, with covariance matrix $\Sigma_\varepsilon$ & mV\\
	$m(\mathbf{r}_n,\mathbf{r}')$ & observation kernel where $n$ is sensor index $n=1,..,n_y$ (width $\sigma_m$) & n.a. \\
	$w(\mathbf{r})$ & spatial connectivity kernel & n.a.\\
	$h(t)$ & post-synaptic response kernel & mV\\
	$\zeta$, $\xi$ & inverse synaptic time constant \& time constant parameter & s$^{-1}$, n.a.\\
	$\eta(t)$ & Heaviside step function & n.a.\\
	$\delta(t)$ & Dirac-delta function & n.a.\\
	$\textrm{D}$ & temporal differential operator & n.a.\\
	\hline
	\multicolumn{3}{|c|}{\emph{Reduced Model}} \\
	\hline
   	$\mathbf{x}_t$ & state vector at time $t$ & mV\\
   	$\mathbf{e}_t$ & state disturbance vector, with covariance $\Sigma_e$ & mV\\
   	$\mathbf{C}$ & observation matrix & n.a. \\
	\hline
	\multicolumn{3}{|c|}{\emph{Frequency Analysis}} \\
	\hline
	$\boldsymbol{\nu}$, $\boldsymbol{\nu}_c$ & spatial frequency and spatial cutoff frequency & cycles/mm \\
	$\sigma_{\nu}^2$ & variance of Fourier transformed Gaussian basis function & (cycles/mm)$^2$\\
	\hline
	\multicolumn{3}{|c|}{\emph{Estimation}} \\
	\hline
	$\mathcal K_{t} $ & filter gain & n.a.\\
	\hline
\end{tabular}
\caption{\textbf{Notation.} The first column of the table gives the symbols for all the notation used in the paper. The second column provides a brief description of the quantity, and the third column provides the SI units where relevant.}
\label{tab:Notation}
\end{table*}
\doublespacing

@@

The state-transition equation can be simplified by exploiting the isotropy of the scaling functions and the wavelets where
\begin{equation}
	\boldsymbol{\lambda}_{j,l}(\mathbf{r}-\mathbf{r}') = \boldsymbol{\lambda}(2\mathbf{c}_{j,l}+\mathbf{r}'-\mathbf{r}). 
\end{equation}
Following this, we can switch the order of operations with the inner product and convolution in equation~\eqref{eq:ApproxDiscreteTimeModel4}. With this switch the convolution is now between the scaling functions and wavelets for the field and the scaling functions and wavelets for the connectivity kernel. This convolution can be calculated analytically when using B-spline wavelets and can be predefined once, since these are static functions. This has major computation advantages when solving the estimation problem. The predefined convolution can be written as
\begin{equation}
	\left[\Gamma(\mathbf{r}')\right]_{:j,l} = T_s \mathbf{\Lambda}_{x}^{-1}\int_{\Omega} \boldsymbol\mu\left(\mathbf{r}\right)\boldsymbol\lambda\left(2\mathbf{c}_{j,l} + \mathbf{r}'-\mathbf{r}\right) \,\mathrm{d}\mathbf{r}
\end{equation}
giving
\begin{equation}
	\mathbf{x}_{t+1} = 
	\xi \mathbf{x}_t + 
	\int_{\Omega} \Gamma\left(\mathbf{r}'\right)f\left(\boldsymbol\mu^\top\left(\mathbf{r}'\right) \mathbf{x}_t\right) 
	\, \mathrm{d}\mathbf{r}' \boldsymbol\theta
	+ \mathbf w_t
\end{equation}
% \begin{align}\label{eq:Lambdax}
%  \mathbf{\Lambda}_{x} &\triangleq \int_{\Omega}\boldsymbol\mu\left(\mathbf{r}\right)\boldsymbol\mu^\top\left(\mathbf{r}\right) \mathrm{d}\mathbf{r} \\
% \label{eq:Lambdatheta}
%  \mathbf{\Lambda}_{\theta} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right) \int_\Omega { 
% 	   \boldsymbol\theta^\top\boldsymbol\lambda\left(r-r'\right)
% 	    \boldsymbol\mu^\top\left(r\right)\ dr'dr.}
% \end{align}
% Substituting \eqref{eq:KernelFiniteExpansion} and \eqref{eq:FieldFiniteExpansion} in \eqref{eq:DiscreteTimeModel}, pre-multiplying by $\boldsymbol\mu\left(r\right)$, and integrating over the space gives
% \begin{align}\label{eq:DecomposedModel2} 
% 	&\mathbf{\Lambda}_{x} \mathbf{x}_{t+1} = 
% 	\xi\mathbf{\Lambda}_{x} \mathbf{x}_t +T_s \varsigma \mathbf{\Lambda}_{\theta}\mathbf{x}_t +\int_{\Omega}\boldsymbol\mu\left(r\right)e_t\left(r\right)dr.
% \end{align}
% Cross-multiplying by $\mathbf{\Lambda}_{x}^{-1}$ gives the state transition equation
% \begin{align}\label{eq:StateEquation}
%  \mathbf x_{t+1} &=\mathbf A(\boldsymbol \theta) \mathbf x_t+ \mathbf w_t\\
% \label{eq:A_theta}
%  \mathbf A(\boldsymbol \theta) &= T_s\varsigma\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda}_{\theta}+\xi\mathbf I.
% \end{align}
% The disturbance becomes 
% \begin{equation}\label{eq:Disturbance}
% \mathbf w_t= \mathbf{\Lambda}_{x}^{-1}\int_{\Omega}\boldsymbol\mu \left(r\right)e_t\left(r\right)dr,
% \end{equation}
% which is a vector valued, zero-mean normally distributed, white noise process with covariance (see \cite{Freestone2011} for proof)
% \begin{equation}\label{eq:CovMatrix}
% \boldsymbol\Sigma_w =\mathbf{\Lambda}_{x}^{-1}\iint\limits_{\boldsymbol\Omega}\boldsymbol\mu\left(r\right) \eta\left(r-r'\right)\boldsymbol\mu^{\top}\left(r'\right)dr'dr\mathbf{\Lambda}_{x}^{-\top}.
% \end{equation}

@@

In order to analytically calculate elements of $\boldsymbol\Lambda_{x}$ and $\boldsymbol\Lambda_{\theta}$, the scaling and wavelet basis functions must be expanded in terms of $N_m$ at the appropriate scale. For scaling functions this can be done  by using \eqref{eq:MRA-TwoScalepair1} recursively, giving
\begin{align}\label{eq:MRA-ConvertingFormulaForscalingFunctions}
 &N_m(r)=\nonumber \\
&\sum_{n_1,n_2, \dots n_j=0}^{m}\left[\mathbf p\right]_{n_1} \left[\mathbf p\right]_{n_2}\dots \left[\mathbf p\right]_{n_j}N_m(2^jr-l_{n_1,n_2, \dots, n_j}),
\end{align}
where 
\begin{align}
 l_{n_1,n_2, \dots, n_j}=2^{j-1}n_1+2^{j-2}n_2+ \dots +2^{0}n_j.
\end{align}
The relation \eqref{eq:MRA-ConvertingFormulaForscalingFunctions} can be also proven by induction (see Appendix ...). A similar method can be used for wavelet functions noting that they must be written first in terms of scaling functions using \eqref{eq:MRA-TwoScalepair2}. Examples of such expansions are given in Figure~\ref{fig:MRA-BasisDecomposition} where the scaling and wavelet functions at $j=0$ are written in terms of 625 and 1250 scaling functions at $j=4$ respectively.
\begin{figure}[!h] 
 	\centering
 		\includegraphics[scale=1]{./Graph/Decomposition.pdf}
 		\caption{{\bf The scaling and the wavelet function decomposition}. (\textbf{A}) The scaling function (blue) at $j=0$ is obtained using 625 scaling functions (black) at $j=4$. (\textbf{B}) The wavelet function (blue) at $j=0$ is obtained using 1250 scaling functions (black) at $j=4$.}
 	\label{fig:MRA-BasisDecomposition}
 \end{figure}  
 In this paper, 4th -- order cardinal B-spline scaling and wavelet functions are used, resulting in  8th and 12th order B-spline functions when calculating \eqref{eq:Lambdax}, \eqref{eq:Lambdatheta} and \eqref{eq:CovMatrix} whose values at integer points are given in Table~\ref{table:MRA-BsplineatIntegerPoints}.
\begin {table}[h]
\begin{center}
\begin{tabular}{lcccc}
 & $k$& $7!N_{8}\left(k\right)$ & $11!N_{12}\left(k\right)$\\ \hline\hline
 & 1  & $1$ & $1$\\
 & 2  & $120$ & $2,036$\\
 & 3  & $1,191$ & $152,637$\\
 & 4  & $2,416$ & $2,203,488$\\
 & 5  & $\space$ & $9,738,114$\\
 & 6  & $\space$ & $15,724,248$
 \end{tabular}
 \caption {{\bf Cardinal B-Splines at the knot sequence}. Table entries are calculated using \eqref{eq:MRA-recursiveBsplineatintegerpoints} and can be also found in \cite{Goswami1999}.} 
 \label{table:MRA-BsplineatIntegerPoints}
 \end{center}
 \end {table}
\subsection{Frequency Analysis}
To determine the level of approximation in the model, detailed in the previous section, that is the number of basis functions required to reconstruct the field, $v_t(r)$, the frequency response of the B-spline function needs to be computed by applying the Fourier transform to \eqref{eq:N1convolutions}. To proceed the Fourier transform of $N_1(r)$ is first calculated, 
\begin{align}\label{eq:MRA-N1Fouriertransform}
\mathcal F(N_1(r))&=\int_{-\infty}^{+\infty}N_1(r)\mathrm{exp}(-2\pi i \nu r)dr \nonumber \\
&=\int_{0}^{1} \mathrm{exp}(-2\pi i \nu r)dr \nonumber \\
&=\frac{1-\mathrm{exp}(-2\pi i \nu)}{2\pi i\nu}.
\end{align}
The convolution theorem states that the Fourier transform of the convolution of functions is equal to the product of their Fourier transforms. Therefore, taking Fourier transform of \eqref{eq:N1convolutions} and substituting \eqref{eq:MRA-N1Fouriertransform} for $\mathcal F(N_1(r)) $ yields
\begin{align}\label{eq:MRA-NmFouriertransform}
\mathcal F(N_m(r))=\left(\frac{1-\mathrm{exp}(-2\pi i \nu)}{2\pi i\nu}\right)^m.
\end{align}
\dean{Need to say how the FT is used somewhere here before proceeding.}

\section{State and Parameter Estimation}
The state-space representation of the MRAIDE allows the use of the well known Expectation Maximization (EM) algorithm \cite{Dempster1977} to infer both the kernel and the field from electrophysiological data. The EM algorithm, when used in this context \cite{Dewar2009}, yields the maximum likelihood kernel estimate and the posterior distribution of the field over time. We use the standard EM algorithm for linear dynamic systems \cite{Shumway2000} and hence only describe aspects of the algorithm that are specific to MRA neural field equations. EM essentially finds increasingly tighter lower bounds on the likelihood of the kernel $p(\mathbf{y}_1 \ldots \mathbf{y}_T|w)$ so that, at convergence, the maximum of the bound corresponds to the maximum of the likelihood. 

@@

\section{Results}\label{sec:MRA-results}
To demonstrate the performance of the MRAIDE estimation framework, data was generated synthetically  using \eqref{eq:DiscreteTimeModel} and \eqref{eq:ObservationEquation}, allowing a comparison between true and estimated parameters. In doing so, the synaptic time constant was set to $\tau = 10$~ms \cite{David2003}, and the sampling time was chosen ten times faster \cite{Stephan2008}, i.e. $T_s = 1$~ms, ensuring the discrepancy between the continuous neural field model and its discrete approximation is small.

Two different experiments were designed where in the first one, 200 runs of one second of data were used, where the field disturbance, $e_t$, and the observation noise, $\boldsymbol\varepsilon_t$, were regenerated each run. This way the net shape of the connectivity kernel could be formed by averaging over the parameters estimates from each realisation. In the second experiment, a single data set was generated and the field estimations are compared for five models, each accounting for a different level of approximation, i.e. $j=0$ to $j=4$.

For each of the experiment, the estimation was applied to the final 900~ms, allowing the model's dynamics to stabilise from the initial conditions. The initial state, parameters and the support of the connectivity kernel were unknown to the estimator. The observation noise was set to $\boldsymbol\Sigma_{\varepsilon}=0.1 \times \mathbf{I}_{n_y}$ and the disturbance covariance function was set to $\gamma(r-r') = \varphi_{3,-2}(r-r')$ (B-spline, $j=3$).  The firing rate slope, $\varsigma$ was set to $0.56~mV^{-1}$ \cite{Wendling2005}. The observation kernel was modelled using a B-spline function, with a width of 0.8~mm at half the maximum amplitude where the distance between adjacent sensors was $0.5$~mm, resulting into $n_y = 161$ observations. The spacing and bandwidth of the sensors allowed the full spatial bandwidth of the field to be observed.  With this number of observations, the estimation problem was well-posed, i.e. the number of states, $n_x < n_y$ at each level of the decomposition. This is because the scaling functions are orthogonal to the wavelets at the same and higher levels of decomposition, and  the wavelets are orthogonal to each other across different scales (see (\ref{eq:MRA-PsiPsiOrthogonality}-\ref{eq:MRA-PhiPsiOrthogonality})). 
\subsection{Experiment I}
The spatial frequency response of the observed field was used to specify the level of decomposition required to represent the field using the wavelets. This is shown in Figure~\ref{fig:ObservationFrequencyResponce} where the graph obtained by averaging the spectral power of the spatial frequency (over time) from the observations \cite{Scerri2009}. The spatial bandwidths of wavelets was calculated analytically using \eqref{eq:MRA-NmFouriertransform}. The result suggested that wavelets up to level $j=3$ (with the bandwidth $\approx[5,8]$ cycles/mm) can represent the significant spatial characteristics of the field which yields $n_x = 131$ states, corresponding to 9 scaling functions at $j=0$ and 8, 16, 32 and 66 wavelet functions respectively at $j=0$, $j=1$, $j=2$ and $j=3$. 
\begin{figure}[!h] 
 	\centering
 		\includegraphics[scale=1]{./Graph/Figure2.pdf}
 		\caption{{\bf Spatial frequency analysis}. The average (over time) power in dB of the spatial frequency of the observations.}
 	\label{fig:ObservationFrequencyResponce}
 \end{figure}

The EM algorithm was allowed to run until the maximum number of 20 iterations was reached, though typically the change in the transition matrix Frobenius norm dropped below $10^{-6}$ after less than 10 iterations. The rate of convergence of the EM based algorithm is depicted in Figure~\ref{fig:MRA-Convergence}. 
\begin{figure}[!h] 
 	\centering
 		\includegraphics[scale=1]{./Graph/ConvergenceRate.pdf}
 		\caption{{\bf Convergence of the EM algorithm}. Representative plot of
the $\mathbf A(\boldsymbol\theta)$ Frobenius norm vs iterations of EM based algorithm. The change in the stopping criterion falls below $10^{-6}$ after less than 10 iterations.}
 	\label{fig:MRA-Convergence}
 \end{figure}     
% The spatial cutoff frequency of the observed field was used to specify the level of decomposition required to represent the field using the wavelets. Using an oversampling parameter of 5, the cutoff spatial frequency was $\nu_{cy} \approx 6.9 $ cycles/mm, obtained by averaging the spectral power of the spatial frequency (over time) from the observations \cite{Scerri2009}. Therefore, wavelets up to level $j=3$ (with the bandwidth $\approx[5,8]$ cycles/mm) can represent the significant spatial characteristics of the field, yielding $n_x = 131$ states.

>>>>

\singlespacing
\begin{table*}[!t]
\begin{tabular}{|l|l|l|}
	\hline
	\textbf{Symbol} & \textbf{Quantity} & \textbf{Units}\\
	\hline
	\multicolumn{3}{|c|}{\emph{Domain and indices}}\\
	\hline
	$\Omega$ & spatial domain & n.a.\\
	$\mathbf{r}$ & spatial location & [mm, mm]\\
	$t$ & time & s\\
	\hline
	\multicolumn{3}{|c|}{\emph{Model}}\\
	\hline
    $y_t$ & observation & mV\\
    $v(\mathbf{r},t)$ & mean membrane potential & mV \\
	$f(v\left(\mathbf{r},t\right))$ & firing rate function & spikes.s$^{-1}$\\
	$e(\mathbf{r},t)$ & field disturbance, with covariance function $\gamma(\mathbf r-\mathbf r')$ & mV\\
	$\boldsymbol\varepsilon_t$ & observation noise, with covariance matrix $\Sigma_\varepsilon$ & mV\\
	$m(\mathbf{r}_n)$ & observation kernel where $n$ is sensor index $n=1,..,n_y$ (width $\sigma_m$) & n.a. \\
	$w(\mathbf{r})$ & spatial connectivity kernel & n.a.\\
	$h(t)$ & post-synaptic response kernel & mV\\
	$\zeta$, $\xi$ & inverse synaptic time constant \& time constant parameter & s$^{-1}$, n.a.\\
	$\eta(t)$ & Heaviside step function & n.a.\\
	$\delta(t)$ & Dirac-delta function & n.a.\\
	$\textrm{D}$ & temporal differential operator & n.a.\\
	\hline
	\multicolumn{3}{|c|}{\emph{Reduced Model}} \\
	\hline
   	$\mathbf{x}_t$ & state vector at time $t$ & mV\\
   	$\mathbf{e}_t$ & state disturbance vector, with covariance $\Sigma_e$ & mV\\
   	$\mathbf{C}$ & observation matrix & n.a. \\
	\hline
	\multicolumn{3}{|c|}{\emph{Frequency Analysis}} \\
	\hline
	$\boldsymbol{\nu}$, $\boldsymbol{\nu}_c$ & spatial frequency and spatial cutoff frequency & cycles/mm \\
	$\sigma_{\nu}^2$ & variance of Fourier transformed Gaussian basis function & (cycles/mm)$^2$\\
	\hline
	\multicolumn{3}{|c|}{\emph{Estimation}} \\
	\hline
	$\mathcal K_{t} $ & filter gain & n.a.\\
	\hline
\end{tabular}
\caption{\textbf{Notation.} The first column of the table gives the symbols for all the notation used in the paper. The second column provides a brief description of the quantity, and the third column provides the SI units where relevant.}
\label{tab:Notation}
\end{table*}
\doublespacing

@@

The state-transition equation can be simplified by exploiting the isotropy of the scaling functions and the wavelets where
\begin{equation}
	\boldsymbol{\lambda}_{j,l}(\mathbf{r}-\mathbf{r}') = \boldsymbol{\lambda}(2\mathbf{c}_{j,l}+\mathbf{r}'-\mathbf{r}). 
\end{equation}
Following this, we can switch the order of operations with the inner product and convolution in equation~\eqref{eq:ApproxDiscreteTimeModel4}. With this switch the convolution is now between the scaling functions and wavelets for the field and the scaling functions and wavelets for the connectivity kernel. This convolution can be calculated analytically when using B-spline wavelets and can be predefined once, since these are static functions. This has major computation advantages when solving the estimation problem. The predefined convolution can be written as
\begin{equation}
	\left[\Gamma(\mathbf{r}')\right]_{:j,l} = T_s \mathbf{\Lambda}_{x}^{-1}\int_{\Omega} \boldsymbol\mu\left(\mathbf{r}\right)\boldsymbol\lambda\left(2\mathbf{c}_{j,l} + \mathbf{r}'-\mathbf{r}\right) \,\mathrm{d}\mathbf{r}
\end{equation}
giving
\begin{equation}
	\mathbf{x}_{t+1} = 
	\xi \mathbf{x}_t + 
	\int_{\Omega} \Gamma\left(\mathbf{r}'\right)f\left(\boldsymbol\mu^\top\left(\mathbf{r}'\right) \mathbf{x}_t\right) 
	\, \mathrm{d}\mathbf{r}' \boldsymbol\theta
	+ \mathbf w_t
\end{equation}

% Substituting \eqref{eq:KernelFiniteExpansion} and \eqref{eq:FieldFiniteExpansion} in \eqref{eq:DiscreteTimeModel}, pre-multiplying by $\boldsymbol\mu\left(r\right)$, and integrating over the space gives
% \begin{align}\label{eq:DecomposedModel2} 
% 	&\mathbf{\Lambda}_{x} \mathbf{x}_{t+1} = 
% 	\xi\mathbf{\Lambda}_{x} \mathbf{x}_t +T_s \varsigma \mathbf{\Lambda}_{\theta}\mathbf{x}_t +\int_{\Omega}\boldsymbol\mu\left(r\right)e_t\left(r\right)dr.
% \end{align}
% Cross-multiplying by $\mathbf{\Lambda}_{x}^{-1}$ gives the state transition equation
% \begin{align}\label{eq:StateEquation}
%  \mathbf x_{t+1} &=\mathbf A(\boldsymbol \theta) \mathbf x_t+ \mathbf w_t\\
% \label{eq:A_theta}
%  \mathbf A(\boldsymbol \theta) &= T_s\varsigma\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda}_{\theta}+\xi\mathbf I.
% \end{align}
% The disturbance becomes 
% \begin{equation}\label{eq:Disturbance}
% \mathbf w_t= \mathbf{\Lambda}_{x}^{-1}\int_{\Omega}\boldsymbol\mu \left(r\right)e_t\left(r\right)dr,
% \end{equation}
% which is a vector valued, zero-mean normally distributed, white noise process with covariance (see \cite{Freestone2011} for proof)
% \begin{equation}\label{eq:CovMatrix}
% \boldsymbol\Sigma_w =\mathbf{\Lambda}_{x}^{-1}\iint\limits_{\boldsymbol\Omega}\boldsymbol\mu\left(r\right) \eta\left(r-r'\right)\boldsymbol\mu^{\top}\left(r'\right)dr'dr\mathbf{\Lambda}_{x}^{-\top}.
% \end{equation}

@@

In order to analytically calculate elements of $\boldsymbol\Lambda_{x}$ and $\boldsymbol\Lambda_{\theta}$, the scaling and wavelet basis functions must be expanded in terms of $N_m$ at the appropriate scale. For scaling functions this can be done  by using \eqref{eq:MRA-TwoScalepair1} recursively, giving
\begin{align}\label{eq:MRA-ConvertingFormulaForscalingFunctions}
 &N_m(r)=\nonumber \\
&\sum_{n_1,n_2, \dots n_j=0}^{m}\left[\mathbf p\right]_{n_1} \left[\mathbf p\right]_{n_2}\dots \left[\mathbf p\right]_{n_j}N_m(2^jr-l_{n_1,n_2, \dots, n_j}),
\end{align}
where 
\begin{align}
 l_{n_1,n_2, \dots, n_j}=2^{j-1}n_1+2^{j-2}n_2+ \dots +2^{0}n_j.
\end{align}
The relation \eqref{eq:MRA-ConvertingFormulaForscalingFunctions} can be also proven by induction (see Appendix ...). A similar method can be used for wavelet functions noting that they must be written first in terms of scaling functions using \eqref{eq:MRA-TwoScalepair2}. Examples of such expansions are given in Figure~\ref{fig:MRA-BasisDecomposition} where the scaling and wavelet functions at $j=0$ are written in terms of 625 and 1250 scaling functions at $j=4$ respectively.
\begin{figure}[!t] 
 	\centering
 		\includegraphics[scale=1]{./Graph/Decomposition.pdf}
 		\caption{{\bf The scaling and the wavelet function decomposition}. (\textbf{A}) The scaling function (blue) at $j=0$ is obtained using 625 scaling functions (black) at $j=4$. (\textbf{B}) The wavelet function (blue) at $j=0$ is obtained using 1250 scaling functions (black) at $j=4$.}
 	\label{fig:MRA-BasisDecomposition}
 \end{figure}  
 In this paper, 4th -- order cardinal B-spline scaling and wavelet functions are used, resulting in  8th and 12th order B-spline functions when calculating \eqref{eq:Lambdax}, \eqref{eq:Lambdatheta} and \eqref{eq:CovMatrix} whose values at integer points are given in Table~\ref{table:MRA-BsplineatIntegerPoints}.
\begin {table}[t]
\begin{center}
\begin{tabular}{lcccc}
 & $k$& $7!N_{8}\left(k\right)$ & $11!N_{12}\left(k\right)$\\ \hline\hline
 & 1  & $1$ & $1$\\
 & 2  & $120$ & $2,036$\\
 & 3  & $1,191$ & $152,637$\\
 & 4  & $2,416$ & $2,203,488$\\
 & 5  & $\space$ & $9,738,114$\\
 & 6  & $\space$ & $15,724,248$
 \end{tabular}
 \caption {{\bf Cardinal B-Splines at the knot sequence}. Table entries are calculated using \eqref{eq:MRA-recursiveBsplineatintegerpoints} and can be also found in \cite{Goswami1999}.} 
 \label{table:MRA-BsplineatIntegerPoints}
 \end{center}
 \end {table}
\subsection{Frequency Analysis}
To determine the level of approximation in the model, detailed in the previous section, that is the number of basis functions required to reconstruct the field, $v_t(r)$, the frequency response of the B-spline function needs to be computed by applying the Fourier transform to \eqref{eq:N1convolutions}. To proceed the Fourier transform of $N_1(r)$ is first calculated, 
\begin{align}\label{eq:MRA-N1Fouriertransform}
\mathcal F(N_1(r))&=\int_{-\infty}^{+\infty}N_1(r)\mathrm{exp}(-2\pi i \nu r)dr \nonumber \\
&=\int_{0}^{1} \mathrm{exp}(-2\pi i \nu r)dr \nonumber \\
&=\frac{1-\mathrm{exp}(-2\pi i \nu)}{2\pi i\nu}.
\end{align}
The convolution theorem states that the Fourier transform of the convolution of functions is equal to the product of their Fourier transforms. Therefore, taking Fourier transform of \eqref{eq:N1convolutions} and substituting \eqref{eq:MRA-N1Fouriertransform} for $\mathcal F(N_1(r)) $ yields
\begin{align}\label{eq:MRA-NmFouriertransform}
\mathcal F(N_m(r))=\left(\frac{1-\mathrm{exp}(-2\pi i \nu)}{2\pi i\nu}\right)^m.
\end{align}
\dean{Need to say how the FT is used somewhere here before proceeding.}

\section{State and Parameter Estimation}
There exist many variations of state and parameter estimation methods for state-space models. The choice of algorithm is normally determined by the dimension of the system, whether the system is linear or nonlinear and whether or not the parameters are considered static or dynamic. Up to this point of the paper, we have not made any assumptions on any of these criteria for choosing an estimation algorithm. The MRAIDE neural field model formulation may be used in any of the above scenarios.

To demonstrate the MRAIDE framework we shall consider the parameters to be static and we shall also assume that the relationship between the mean membrane potential and the mean firing rate is predominantly linear. Note that there is a difference between forward and inverse modelling when making these assumptions. For example, by using a linear activation function in the estimator we are not assuming that there is necessarily a linear relationship between the membrane voltage and firing rate, but that the statistics of the neural masses fall within a linear operating region over the estimation time period.

It is important to point out there is no doubt that there will always be a mismatch between neural mass models and actual cortical tissue. Nevertheless, the simplifications used in formulating the equation enable the development of new methods for extracting information from electrophysiological data, that has the potential to influence the treatment of diseases such as epilepsy. The trade-off between parsimony and complexity is always difficult when considering neurodynamics and this is further complicated when using estimation algorithms that have huge computational demands with such high-dimensional systems.

By assuming a linear activation function of the form
\begin{equation}
	f(v_t(\mathbf{r})) = \varsigma v_t(\mathbf{r})
\end{equation}
the state transition equation~\eqref{eq:ApproxDiscreteTimeModel4} simplifies to
\begin{equation}\label{eq:ApproxDiscreteTimeModel_Linear}
	\mathbf{x}_{t+1} = 
	\xi \mathbf{x}_t + 
	\varsigma T_s \mathbf{\Lambda}_{x}^{-1} \int_{\Omega}\boldsymbol\mu\left(\mathbf{r}\right)\int_\Omega { 
	    \boldsymbol\theta^\top\boldsymbol\lambda\left(\mathbf{r}-\mathbf{r}'\right)
	    \boldsymbol\mu^\top\left(\mathbf{r}'\right) 
	\, \mathrm{d}\mathbf{r}'\mathrm{d}\mathbf{r}} \mathbf{x}_t
	+ \mathbf{w}_t.
\end{equation}
Now we can further simplify the system by defining
\begin{align}
	\label{eq:Lambdatheta}
	 \mathbf{\Lambda}_{\theta} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right) \int_\Omega { 
		   \boldsymbol\theta^\top\boldsymbol\lambda\left(r-r'\right)
		    \boldsymbol\mu^\top\left(r\right)\ \mathrm{d}r'\mathrm{d}r}
\end{align}
to give the state transition equation
\begin{align}\label{eq:StateEquation}
 \mathbf x_{t+1} &=\mathbf A(\boldsymbol \theta) \mathbf x_t+ \mathbf w_t\\
\label{eq:A_theta}
 \mathbf A(\boldsymbol \theta) &= T_s\varsigma\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda}_{\theta}+\xi\mathbf I.
\end{align}

The linear state-space representation of the MRAIDE allows the use of the well known Expectation Maximisation (EM) algorithm \cite{Dempster1977} to infer both the kernel and the field from electrophysiological data. The EM algorithm, when used in this context \cite{Dewar2009}, yields the maximum likelihood kernel estimate and the posterior distribution of the field over time. We use the standard EM algorithm for linear dynamic systems \cite{Shumway2000} and hence only describe aspects of the algorithm that are specific to MRA neural field equations. EM essentially finds increasingly tighter lower bounds on the likelihood of the kernel $p(\mathbf{y}_1 \ldots \mathbf{y}_T|w)$ so that, at convergence, the maximum of the bound corresponds to the maximum of the likelihood. 

@@

\section{Results}\label{sec:MRA-results}
To demonstrate the performance of the MRAIDE estimation framework, data was generated synthetically  using equations \eqref{eq:DiscreteTimeModel} and \eqref{eq:ObservationEquation}, allowing a comparison between true and estimated parameters. In doing so, the membrane time constant was set to $\tau = 10$~ms \cite{David2003}, and the sampling time was chosen ten times faster \cite{Stephan2008}, i.e. $T_s = 1$~ms, ensuring the discrepancy between the continuous neural field model and its discrete approximation is small.

Two different experiments were designed to test the framework. In the first one, 200 realisations of 1 second of data were used, where the field disturbance, $e_t$, and the observation noise, $\boldsymbol\varepsilon_t$, were regenerated each run. This way the net shape of the connectivity kernel could be formed by averaging over the parameters estimates from each realisation. In the second experiment, a single data set was generated and the field estimations are compared for five models, each accounting for a different level of approximation, i.e. $j=0$ to $j=4$.

For each of the experiments, the estimation was applied to the final 900~ms, allowing the forward model's dynamics to stabilise from the initial conditions. The initial state, parameters and the support of the connectivity kernel were unknown to the estimator. The observation noise was set to $\boldsymbol\Sigma_{\varepsilon}=0.1 \times \mathbf{I}_{n_y}$ and the disturbance covariance function was set to $\gamma(r-r') = \varphi_{3,-2}(r-r')$ (B-spline, $j=3$).  The firing rate slope, $\varsigma$ was set to $0.56~mV^{-1}$ \cite{Wendling2005}. The observation kernel was modelled using a B-spline function, with a width of 0.8~mm at half the maximum amplitude where the distance between adjacent sensors was $0.5$~mm, resulting into $n_y = 161$ observations. The spacing and bandwidth of the sensors allowed the full spatial bandwidth of the field to be observed.  With this number of observations, the estimation problem was well-posed, i.e. the number of states, $n_x < n_y$ at each level of the decomposition. This is because the scaling functions are orthogonal to the wavelets at the same and higher levels of decomposition, and  the wavelets are orthogonal to each other across different scales (see (\ref{eq:MRA-PsiPsiOrthogonality}-\ref{eq:MRA-PhiPsiOrthogonality})).

\subsection{Experiment I}
The spatial frequency response of the observed field was used to specify the level of decomposition required to represent the field using the wavelets. This is shown in Figure~\ref{fig:ObservationFrequencyResponce} where the graph obtained by averaging the spectral power of the spatial frequency (over time) from the observations \cite{Scerri2009}. The spatial bandwidths of wavelets was calculated analytically using equation \eqref{eq:MRA-NmFouriertransform}. The result suggested that wavelets up to level $j=3$ (with the bandwidth $\approx[5,8]$ cycles/mm) can represent the significant spatial characteristics of the field which yields $n_x = 131$ states, corresponding to 9 scaling functions at $j=0$ and 8, 16, 32 and 66 wavelet functions, respectively at $j=0$, $j=1$, $j=2$ and $j=3$. 
\begin{figure}[!t] 
 	\centering
 		\includegraphics[scale=1]{./Graph/Figure2.pdf}
 		\caption{{\bf Spatial frequency analysis}. The average (over time) power in dB of the spatial frequency of the observations.}
 	\label{fig:ObservationFrequencyResponce}
 \end{figure}

The EM algorithm was allowed to run until a maximum number of 20 iterations was reached, though typically the change in the transition matrix Frobenius norm dropped below $10^{-6}$ after less than 10 iterations. The rate of convergence of the EM based algorithm is depicted in Figure~\ref{fig:MRA-Convergence}. 
\begin{figure}[!t] 
 	\centering
 		\includegraphics[scale=1]{./Graph/ConvergenceRate.pdf}
 		\caption{{\bf Convergence of the EM algorithm}. Representative plot of
the $\mathbf A(\boldsymbol\theta)$ Frobenius norm vs iterations of EM based algorithm. The change in the stopping criterion falls below $10^{-6}$ after less than 10 iterations.}
 	\label{fig:MRA-Convergence}
 \end{figure}     
% The spatial cutoff frequency of the observed field was used to specify the level of decomposition required to represent the field using the wavelets. Using an oversampling parameter of 5, the cutoff spatial frequency was $\nu_{cy} \approx 6.9 $ cycles/mm, obtained by averaging the spectral power of the spatial frequency (over time) from the observations \cite{Scerri2009}. Therefore, wavelets up to level $j=3$ (with the bandwidth $\approx[5,8]$ cycles/mm) can represent the significant spatial characteristics of the field, yielding $n_x = 131$ states.
